### 2022-10-14 draft.tex
Deploying to gh-pages from @ gottacatchenall/ms_false_negatives@230beaa0261730b22ad4cadb32c74b1331cb84d4 🚀

Understanding which and how species interact underlies many questions in
evolutionary biology and ecology, but also is an increasing imperative
to both mitigate the anthropogenic change on Earth's biodiversity
(Jordano 2016a; Makiola \emph{et al.} 2020) and to predict potential
spillover of zoonotic disease (Becker \emph{et al.} 2021). Over the past
few decades biodiversity data has become increasingly available:
remote-sensing has enabled collection of data on spatial scales and
resolutions previously unimaginable, automation of in-situ sensing
(Stephenson 2020), and adoption of open data practices (Kenall \emph{et
al.} 2014) have substantially amount of data available to ecologists.
Still, widespread data about species \emph{interactions} remains elusive
(\textbf{TIM\_DATASET\_PAPER?}). Often, observing an interaction between
two species requires human sampling, typically because remote sampling
methods can only detect co-occurrence, and this itself is not
necessarily indicative of interaction (Blanchet \emph{et al.} 2020).
This constraint induces biases on species interaction data subject to
the spatial and temporal scales that humans can feasibly sample.

\emph{Sampling effort} and its impact on the resulting data collected
from ecosystems has encouraged a long history of discourse. The recorded
number of species in a sample depends on the total number of
observations (Walther \emph{et al.} 1995; Willott 2001), as do estimates
of population abundance (Griffiths 1998). This has motivated more
quantitatively robust approaches to account for error in sampling data
in many contexts: to determine if a given species is extinct (Boakes
\emph{et al.} 2015), to determine sampling design (Moore \& McCarthy
2016), and to measure species richness across large scales (Carlson
\emph{et al.} 2020). In the context of interactions, the initial concern
was the compounding effects of limited sampling effort combined with the
amalgamation of data (across both study sites and across taxonomic
scales) could lead any empirical set of observations to inadequately
reflect the reality of how species interact (Paine 1988). Martinez
\emph{et al.} (1999) showed that in a plant-endophyte trophic network,
network connectance is robust to sampling effort, but this done in the
context of a system for which observation of 62,000 total interactions
derived from 164,000 plant-stems was feasible. In some systems
(e.g.~megafauna food-webs) this many observations is either impractical
or infeasible due to the absolute abundance of the species in question.

Because we cannot feasibly observe all (or even many) of the
interactions that occur in nature, our samples end up capturing only a
small fraction of those interactions. This means we can be reasonably
confident two species actually interact if we have a record of it, but
not at all confident that two species \emph{do not} interact if we have
no record of those species observed together. In other words, it is
difficult to distinguish true-negatives (two species \emph{never}
interact) from \emph{false-negatives} (two species interact in some
capacity, but we have not observed it). This is then amplified as the
interaction data we have is geographically toward the usual suspects
(Poisot \emph{et al.} 2021a), This noise in data has practical
consequences for answering questions about species interactions (de
Aguiar \emph{et al.} 2019)---these false-negatives could go on to effect
the inferences we make about network properties and relations among
species, and our predictions about how species will interact in the
future.

This is compounded by semantic confusion about the definition of
``interaction''. Here distinguish between: a species \emph{occurring}, a
species being \emph{observed occurring}, two species being observed
\emph{co-occurring}, and two species being observed \emph{interacting}
(fig.~\ref{fig:taxonomy}). In this manuscript, we refer to species
either as ``interacting''---two species co-occur (and, at least
sometimes, interact)---or ``not-interacting'' (two species that,
regardless of whether they co-occur, neither exhibits any meaningful
effect on the biomass of the other). In fig.~\ref{fig:taxonomy} we see
that, under our definition, observing two species co-occurring is a
prerequisite for observing an interaction between two species.

But species are not observed with equal probability but instead in
proportion to their relative biomass---you are much more likely to
observe a species of high relative abundance than one of very low
relative abundance (Poisot \emph{et al.} 2015). This assumes that there
are no associations in species co-occurrence due to an interaction
(perhaps because this interaction is ``important'' for both species)
(Cazelles \emph{et al.} 2016), but here we show increasing strength of
associations leads to increasing probability of false-negatives in
interaction data. Further observed co-occurrence is often equated with
meaningful interaction strength, but this is not necessarily the case
(Blanchet \emph{et al.} 2020; Strydom \emph{et al.} 2021). Bears and
salmon \emph{interact}---a bear and the microbes in the soil of a dens
interact, but less so.

Here, we show that the probability of observing a actual
``non-interaction'' between species depends on sampling effort, and
suggest that surveys of species interactions can benefit from simulation
modeling of detection probability (Jordano 2016b). We demonstrate that
the realized false-negative rate of interactions is directly related the
relative abundance of a particular species, relationship between total
sampling effort (the total count of all individuals of all species seen)
and false-negative rate. questions we pose and attempt to answer are: 1)
How many times do you have to observe a non-interaction between two
species to be confident in saying that is a true negative? 2) How
``wrong'' are the measurements of network structure as a function of
false-negative probability? and lastly 3) How do false-negatives impact
our ability to make reliable predictions about interactions? We show
that positive associations in co-occurrence data can increase realized
probability of false negatives, and demonstrate these positive
associations are present in two spatially-replicated systems. We
conclude by suggesting that simulation of sampling effort and species
occurrence can and should be used to help design surveys of species
diversity (Moore \& McCarthy 2016), and by advocating use of null models
like those presented here as a tool for guiding design of surveys of
species interactions, and for modeling detection error in predictive
ecological models.

\hypertarget{how-many-observations-of-a-non-interaction-do-we-need-to-classify-it-as-a-true-negative}{%
\section{How many observations of a non-interaction do we need to
classify it as a true
negative?}\label{how-many-observations-of-a-non-interaction-do-we-need-to-classify-it-as-a-true-negative}}

To answer the titular question of this section, we present a naive model
of interaction detection: we assume that every interacting pair of
species is incorrectly observed as a not-interacting with an independent
and fixed probability, which we denote \(p_{fn}\) and subsequently refer
to as the False-Negative Rate (FNR). If we observe the same species
not-interacting \(N\) times, then the probability of a true-negative
(denoted \(p_{tn}\)) is given by \(p_{tn} = 1 - (p_{fn})^N\). This
relation (callend the geometric distribution, a special case of the
negative-binomial distribution) is shown in fig.~\ref{fig:negativebinom}
for varying values of the false negative rate \(p_{fn}\). This
illustrates a fundamental link between our ability to reliably say an
interaction doesn't exist---\(p_{tn}\)---and the number of times we have
observed a given species. In addition, note that there also is no
non-zero \(p_{fn}\) for which we can ever \emph{prove} that an
interaction does not exist---no matter how many observations of
non-interaction \(N\) we have, \(p_{tn} < 1\).

From fig.~\ref{fig:fig1} (A) (and general intuition) it is clear that
the more times we see two species \emph{occurring}, but \emph{not}
interacting, the more likely the interaction is a true negative. But how
does one decide what this threshold of number of observations should be
when planning to sample a given system? If false-negative rates
presented in fig.~\ref{fig:fig1} seem unrealistically high, consider
that species are not observed independent of their relative abundance.
In the next section we demonstrate that distribution of abundance in
ecosystems can lead to realized values of \(p_{fn}\) similar to those in
fig.~\ref{fig:negativebinom} for species with low relative abundance,
simply as a artifact of sampling effort.

\hypertarget{false-negatives-as-a-product-of-relative-abundance}{%
\subsection{False-negatives as a product of relative
abundance}\label{false-negatives-as-a-product-of-relative-abundance}}

Here we show the realized false-negative rate of species interactions
changes drastically with sampling effort, largely due to the intrinsic
variation of abundances within a community. We do this by simulating the
process of observation of species interactions, applied both to 243
empirical food webs from the Mangal database (Banville \emph{et al.}
2021) as well as random food-webs generated using the niche model
(Williams \& Martinez 2000). Our neutral model of observation assumes
each observed species is drawn from the distribution of those species'
abundances at that place and time. Although there is no shortage of
debate as to the processes the govern this distribution of abundances
within a community, this abundance distribution can be reasonably-well
described by a log-normal distribution (Volkov \emph{et al.} 2003) (Note
that in addition to the log-normal distribution, we also tested the case
where the abundance distribution is derived from power-law scaling
\(Z^{(T_i-1)}\) where \(T_i\) is the trophic level of species \(i\) and
\(Z\) is a scaling coefficient. (Savage \emph{et al.} 2004), which
yields the same qualitative behavior, \emph{supplement figure 1}). The
practical consequence of abundance distributions spanning many order of
magnitude is seeing two ``rare'' species interacting requires two low
probability events: observing two rare species \emph{at the same time}.

To simulate the process of observation, for an ecological network \(A\)
with \(S\) species, we sample abundances for each species from a
standard-log-normal distribution. For each true interaction in \(A\)
(i.e.~\(A_{ij} = 1\)) we estimate the probability of observing both
species \(i\) and \(j\) at given place and time by simulating \(n\)
observations of individuals, where the species of the individual
observed at the \(1,2,\dots,n\)-th observation is drawn from the
generated log-normal distribution of abundances. For each pair of
species \((i,j)\), if both \(i\) and \(j\) are observed within the \(n\)
observations, the interaction is tallied as a true positive if
\(A_{ij}=1\) and a false positive otherwise. Similarly, if only one of
\(i\) and \(j\) are observed---\emph{but not both}---in these \(n\)
observations, but \(A_{ij}=1\), this is counted as a false-negative, and
a true-negative otherwise.

In fig.~\ref{fig:fig1} (C) and (D) we see this model of observation
applied to networks generated using the niche model (Williams \&
Martinez 2000) across varying levels of species richness, and in (b)
applied to 243 food-webs from the Mangal database. For all niche model
simulations in this manuscript, for a given number of species \(S\) the
number of interactions is drawn from the flexible-links model fit to
Mangal data (MacDonald \emph{et al.} 2020), effectively drawing the
number of interactions \(L\) for a random niche model food-web as
\(L \sim \text{BetaBinomial}(S^2-S+1, \mu \phi, (1-\mu)\phi)\), where
the MAP estimate of (\(\mu\), \(\phi\)) applied to Mangal data from
MacDonald \emph{et al.} (2020) is \((\mu = 0.086, \phi =24.3)\). All
simulations were done with 500 independent replicates per unique number
of observations \(n\). All analyses presented here are done in Julia
v1.6 (Bezanson \emph{et al.} 2015) using both EcologicalNetworks.jl v0.5
and Mangal.jl v0.4 {[}Banville \emph{et al.} (2021); ZENODO link
TODO{]}. Note that the empirical data also is, due to the phenomena
described here, very likely to \emph{already} have many false negatives,
which is why we are interested in prediction of networks in the first
place---we'll revisit this in the final section.

\begin{figure}
\hypertarget{fig:fig1}{%
\centering
\includegraphics{./figures/fig1.png}
\caption{A) The probability an observed interaction is a true negative
(y-axis) given how many times it has been sampled as a non-interaction
(x-axis). Each color reflects a different value of \(p_{fn}\), the
false-negative rate (FNR). this is effectively the cdf of the
negative-binomial distribution with \(r=1\). (B) The expected needed
observations of all individuals of all species (y-axis) required to
obtain a goal number of observations (colors) of a particular species,
and a function of the relative abundance of that focal species (x-axis).
(C) and (D): False negative rate (y-axis) as a function of total
sampling effort (x-axis) and network size, computed using the method
described above. For 500 independent draws from the niche model
Williams2000SimRul at varying levels of species richness (colors) with
connectance drawn according to the flexible-links model
MacDonald2020RevLin as described in the main text. For each draw from
the niche model, 200 sets of 1500 observations are simulated, for which
each the mean false negative rate at each observation-step is computed.
Means denoted with points, with \(1\sigma\) in the first shade and
\(2\sigma\) in the second. B: empirical food webs from Mangal database
in teal, applied to the same process as the A. The outlier on panel B is
a 714 species food-web}\label{fig:fig1}
}
\end{figure}

In panel (c) of fig.~\ref{fig:totalobs}, we show the expected number of
total observations needed to obtain a ``goal'' number of observations
(colors) of a particular ``focal'' species. As an example, if we
hypothesize that \(A\) and \(B\) do not interact, and we want to see
species \(A\) and \(B\) both co-occurring and not-interacting 10 times
to be confident this is a negative (a la fig.~\ref{fig:negativebinom}),
then we need an expected 10,000 observations of all species if the
relative abundance of \(A\) is 0.00125.

Empirical data on interactions are subject to the practical limitations
of funding and human-work hours, and therefore existing data tend to
fall on the order on 100s or 1000s observations of individuals per site
(Nielsen \& Bascompte 2007; Schwarz \emph{et al.} 2020; Resasco \emph{et
al.} 2021). Clear aggregation of this data has proven difficult to find
and a meta-analysis of network data and sampling effort seems both
pertinent and necessary, in addition to the effects of aggregation of
interactions across taxonomic scales (Gauzens \emph{et al.} 2013;
Giacomuzzo \& Jordán 2021). Further, from fig.~\ref{fig:totalobs} it is
evident that the number of species considered in a study is inseparable
from the false-negative rate in that study, and this effect should be
taken into account when designing samples of ecological networks in the
future.

We conclude this section by advocating for the use of neutral models
similar to above to generate expectations about the number of
false-negatives in a data set of a given size. This could prove fruitful
both for designing surveys of interactions (Canard \emph{et al.} 2012),
but also because we may want to incorporate models of observation error
into predictive models (Joseph 2020). Additionally, one must consider
the context for sampling---is the goal to detect a particular species
\(A\) (as in fig.~\ref{fig:totalobs} (c)), or to get a representative
sample of interactions across the species pool? This argument is
well-considered when sampling species (Willott 2001), but has not yet
been internalized for designing samples of communities.

\hypertarget{positive-associations-can-increase-the-probability-of-false-negatives}{%
\subsection{Positive associations can increase the probability of
false-negatives}\label{positive-associations-can-increase-the-probability-of-false-negatives}}

This model above doesn't consider the possibility that there are
positive or negative associations which shift the probability of
observing two species together due to their interaction (Cazelles
\emph{et al.} 2016). However, here we demonstrate that the probability
of observing a false negative can be \emph{higher} if there is some
positive association between occurrence of species \(A\) and \(B\).

If we denote the probability that we observe an interaction we know
exists between \(A\) and \(B\) as \(P(AB)\), and if there is \emph{no}
association between the marginal probabilities of observing \(A\) and
observing \(B\), denoted \(P(A)\) and \(P(B)\) respectively, then the
probability of observing the interaction \(P(AB) = P(A)P(B)\). In the
other case where there \emph{is} some positive strength of association
between observing both \(A\) and \(B\) because this interaction is
``important'' for each species, then the probability of observation both
\(A\) and \(B\), \(P(AB)\), is greater than \(P(A)P(B)\) as \(P(A)\) and
\(P(B)\) are not independent and instead are positively correlated,
\emph{i.e.} \(P(AB) > P(A)P(B)\). In this case, the probability of
observing a false negative in our naive model from
fig.~\ref{fig:negativebinom} is \(p_{fn} = 1 - P(AB)\) which due to the
above inequality implies \(p_{fn} \geq 1 - P(A)P(B)\) which indicates
increasingly greater probability of a false negative as
\(P(AB) \to P(AB) \gg P(A)P(B)\).

However this does not consider variation in species abundance in space
and time, (Poisot \emph{et al.} 2015). If positive or negative
associations between species structure variation in the distribution of
\(P(AB)\) across space/time, then the spatial/temporal biases induced by
data collection would further impact the realized false negative rate,
as the probability of false negative would not be constant for each pair
of species across sites. To test for this association empirical data, we
use two datasets: a set of host-parasite interactions sampled across 51
sites with 327 total taxa (Hadfield \emph{et al.} 2014) and a set of 18
New Zealand freshwater stream food webs with 566 total taxa (Thompson \&
Townsend 2000). We simply compute the empirical marginal distribution of
species occurrence, and compare the product of the marginals,
\(P(A)P(B)\), to the empirical joint distribution \(P(AB)\).

\begin{figure}
\hypertarget{fig:associations}{%
\centering
\includegraphics{./figures/positiveassociations.png}
\caption{Top: Hadfield, Bottom: NZ Stream Foodwebs. Effectively a
version of Cazelles \emph{et al.} (2016) figure 1 panel A. Both
distributions have \(\mu \neq 0\) with
\(p < 10^{-50}\)}\label{fig:associations}
}
\end{figure}

In fig.~\ref{fig:associations}, both host-parasite system (top) and
food-web (bottom) exhibit these positive associations. There is no
reason to expect the strength of this association to be the same in
different systems. At the moment, computing this metric for all of the
networks in the Mangal database proves challenging as most data sets use
different taxonmic identifiers, often at different resolutions. These
particular datasets (Thompson \& Townsend 2000; Hadfield \emph{et al.}
2014) were usable because they already have been sorted to have a fixed
taxonomic backbone (as part of EcologicalNetworks.jl (Banville \emph{et
al.} 2021)). Applying this in bulk to Mangal food-webs presents the
difficulty of resolving different taxon identifiers across spatial
samples of species with to different resolutions, which is why we can't
simply apply this to the whole Mangal database---this highlights a
general problem of resolving taxonomic indentifiers which use different
names and different resolutions in different ecological datasets, which
is a problem that needs to be addressed for computational approaches to
scale up to the world of big-ecological-data we hope to build, although
this is a task that may be aided via natural-language-processing
methods.

\hypertarget{the-impact-of-false-negatives-on-network-analysis-and-prediction}{%
\section{The impact of false-negatives on network analysis and
prediction}\label{the-impact-of-false-negatives-on-network-analysis-and-prediction}}

We now transition toward assessing the effects of false negatives in
data on the properties of the networks which we derive from this
interaction data, and their effect on models for predicting interactions
in the future.

\hypertarget{effects-of-false-negatives-on-network-properties}{%
\subsection{Effects of false-negatives on network
properties}\label{effects-of-false-negatives-on-network-properties}}

>>>>

Species interactions drive many processes in evolutionary biology and
community ecology. A better understanding of interactions among species
is an imperative to both mitigate the potentially harmful impacts of
anthropogenic change on Earth's biodiversity (Makiola \emph{et al.}
2020) and to predict zoonotic spillover of disease to prevent future
pandemics (Becker \emph{et al.} 2021). However, meeting these challenges
is difficult because interactions are intrinsically hard to sample
(Jordano 2016). Over the past few decades biodiversity data has become
increasingly available--for example, remote-sensing has enabled
collection of data on spatial scales and resolutions previously
unimaginable, and improved in-situ sensing (Stephenson 2020), while the
adoption of open data practices (Kenall \emph{et al.} 2014) have
substantially increased the amount of data available to ecologists.
Still, widespread data about species interactions remains elusive
(Poisot \emph{et al.} 2021). Observing an interaction between two
species often requires human observation because remote sampling methods
can primarily detect co-occurrence (Niedballa \emph{et al.} 2019), and
this itself is not necessarily indicative of an interaction (Blanchet
\emph{et al.} 2020). This constraint induces biases on species
interaction data subject to the spatial and temporal scales that current
observation methods can feasibly sample. This is further compounded by
semantic confusion around the word ``interaction''---for example one
could consider competition a form of species interaction, even though
this is marked by a lack of co-occurrence between species, unlike other
types of interactions, like trophism or pollination, which require both
species to be together at the same place and time. Here we define
interaction in the latter sense, where two species have fitness
consequences on one-another if they are in the sample place at the same
time. Further we only consider direct (not higher-order) interactions.

The importance of sampling effort and its impact on resulting ecological
data has produced a rich body of literature. The recorded number of
species in a dataset or sample depends on the total number of
observations (Walther \emph{et al.} 1995; Willott 2001)---as do
estimates of population abundance (Griffiths 1998)---in addition to
spatial coverage and species detectability. This has motivated more
quantitatively robust approaches to account for error in sampling data
in many contexts: to determine if a given species is extinct (Boakes
\emph{et al.} 2015), to determine sampling design (Moore \& McCarthy
2016),and to measure species richness across large scales (Carlson
\emph{et al.} 2020). In the context of interactions, an initial concern
was the compounding effects of limited sampling effort combined with the
amalgamation of data (across both study sites, time of year, and
taxonomic scales) could lead any empirical set of observations to
inadequately reflect the reality of how species interact
{[}Paine1988RoaMap{]} or the structure of the network as a whole
{[}McLeod2021SamAsy{]}. Martinez \emph{et al.} (1999) showed that in a
plant-endophyte trophic network, network connectance is robust to
sampling effort, but this was done in the context of a system for which
observation of 62,000 total interactions derived from 164,000
plant-stems was feasible. In some systems (e.g.~megafauna food-webs)
this many observations is either impractical or infeasible due to the
absolute abundance of the species in question.

We cannot feasibly observe all (or even most) of the direct interactions
that occur in an ecosystem. This means we can be confident two species
actually interact if we have a record of it (given an estimate of
species misidentification probability), but not at all confident that a
pair of species do not interact if we have no record of those species
observed together. In other words, it is difficult to distinguish
true-negatives (two species never interact) from false-negatives (two
species interact in some capacity, but we do not have a record of it).
For a concrete example of this phenomenon, see Figure Concept. Because
even the most highly sampled systems will still contain missing
interactions, there is increasing interest in combining species-level
data (e.g. traits, abundance, range, phylogenetic relatedness, etc.) to
build predictive models of interactions (Strydom \emph{et al.} 2021).
However, the noise of false-negatives could impact the efficacy of our
predictive models and have practical consequences for answering
questions about interactions (de Aguiar \emph{et al.} 2019). This data
constraint is amplified as the interaction data we have is
geographically biased toward the usual suspects (Poisot \emph{et al.}
2021). We therefore need a systematic approach to assessing these biases
in the observation process and the consequences this has for our
understanding of interaction networks

The intrinsic properties of ecological communities create several
challenges for sampling: first, species are not observed with equal
probability---we are much more likely to observe a species of high
abundance than one of very low abundance (Poisot \emph{et al.} 2015).
Canard \emph{et al.} (2012) presents a null model of food-web structure
where species encounter one-another directly in proportion to each
species' relative-abundance. This assumes that there are no associations
in species co-occurrence due to an interaction (perhaps because this
interaction is ``important'' for both species; Cazelles \emph{et al.}
(2016)), but in this paper we later show increasing strength of
associations leads to increasing probability of false-negatives in
interaction data, and that these positive associations are rampant in
existing network data. Second, observed co-occurrence is often equated
with meaningful interaction strength, but this is not necessarily the
case Blanchet \emph{et al.} (2020)---a true ``non-interaction'' would
require that neither of two species, regardless of whether they
co-occur, exhibit any meaningful effect on the fitness of the other. So,
although co-occurrence is not directly indicative of an interaction, it
is a precondition for interactions, and therefore observations of
``non-interactions'' between pairs of species that are outside of the
union of both species ranges do not provide any information about that
interaction, i.e.~they should be excluded from consideration.

Here, we illustrate how our confidence that a pair of species we believe
to not interact highly depends on sampling effort, and suggest that
surveys of species interactions can benefit from simulation modeling of
detection probability. We demonstrate that the realized false-negative
rate of interactions is directly related to the relative abundance of
the species involved, and demonstrate how simulation can be used to
produce a null estimate of the false-negative rate as a function of
total sampling effort (the total count of all individuals of all species
seen). We show that positive associations in co-occurrence data can
increase realized probability of false negatives, and demonstrate these
positive associations are ubiquitous in network datasets. We conclude by
recommending that the simulation of sampling effort and species
occurrence can and should be used to help design surveys of species
interaction diversity (Moore \& McCarthy 2016), and by advocating use of
null models like those presented here as a tool for both guiding design
of surveys of species interactions and for modeling detection error in
predictive models.

\begin{figure}
\hypertarget{fig:concept}{%
\centering
\includegraphics{./figures/concept.png}
\caption{This conceptual example considers a sample of the trophic
community of bears, wolves, salmon (pink fish), pike (yellow fish),
berry trees, and aspen trees. The true metaweb (all realized
interactions across the entire spatial extent) is shown on the left. In
the center is what a hypothetical ecologist samples at each site. Notice
that although bears are observed co-occurring with both salmon and pike,
there was never a direct observation of bears eating pike, even though
they actually do. Therefore, this interaction between bears and pike is
a false negative.}\label{fig:concept}
}
\end{figure}

\hypertarget{how-many-observations-of-a-non-interaction-do-we-need-to-be-confident-its-a-true-negative}{%
\section{How many observations of a non-interaction do we need to be
confident it's a true
negative?}\label{how-many-observations-of-a-non-interaction-do-we-need-to-be-confident-its-a-true-negative}}

We start with a naive model of interaction detection: we assume that
every interacting pair of species is incorrectly observed as a
not-interacting with an independent and fixed probability, which we
denote pfn and subsequently refer to as the False-Negative Rate (FNR).
If we observe the same species not-interacting N times, then the
probability of a true-negative (denoted \(p_{tn}\)) is given by
\(p_{tn}=1-(p_{fn})^N\). This relation (the probability-mass-function of
geometric distribution, a special case of the negative-binomial
distribution) is shown in Figure 1(A) for varying values of \(p_{fn}\)
and illustrates a fundamental link between our ability to reliably say
an interaction doesn't exist---\(p_{tn}\)---and the number of times we
have observed a given species. In addition, note that also there is no
non-zero \(p_{fn}\) for which we can ever prove that an interaction does
not exist---no matter how many observations of non-interactions \(N\) we
have, \(p_{tn}<1\).

From Figure 1(A) it is clear that the more often we see two species
co-occurring, but not interacting, the more likely the interaction is a
true negative. This has several practical consequences: first it means
negatives taken outside the overlap of the range of each species aren't
informative. Second, we can use this relation to compute the expected
number of total observations needed to obtain a ``goal'' number of
observations of a particular pair of species (Figure 1(B)). As an
example, if we hypothesize that \(A\) and \(B\) do not interact, and we
want to see species \(A\) and \(B\) both co-occurring and not
interacting 10 times to be confident this is a true negative, then we
need an expected 1000 observations of all species if the relative
abundances of \(A\) and \(B\) are both \(0.1\).

Because the true FNR is latent, we can never actually be sure what the
actual number of false negatives in our data---however, we can use
simulation to estimate it for datasets of a given size using neutral
models of observation. If some of the ``worst-case'' FNRs presented in
fig.~\ref{fig:geometric}(A) seem unrealistically high, consider that
species are observed in proportion to their relative abundance. In the
next section we demonstrate that the distribution of abundance in
ecosystems can lead to very high realized values of FNR (\(p_{fn}\))
simply as an artifact of sampling effort.

\begin{figure}
\hypertarget{fig:geometric}{%
\centering
\includegraphics{./figures/fig1.png}
\caption{(A) The probability that an observed interaction is a true
negative (y-axis) given how many times it has been sampled as a
non-interaction (x-axis). Each color reflects a different value of
\(p_{fn}\), the false-negative rate (FNR)---this is effectively the cdf
of the geometric distribution. (B): The expected needed observations of
all individuals of all species (y-axis) required to obtain a goal number
of observations (colors) of a particular species, and a function of the
relative abundance of that focal species (x-axis). (C): False negative
rate (y-axis) as a function of total sampling effort (x-axis) and
network size, computed using the method described above. For 500
independent draws from the niche model (Williams \& Martinez (2000)) at
varying levels of species richness (colors) with connectance drawn
according to the flexible-links model (MacDonald \emph{et al.} (2020))
as described in the main text. For each draw from the niche model, 200
sets of 1500 observations are simulated, for which each the mean false
negative rate at each observation-step is computed. Means denoted with
points, with 1 in the first shade and 2 in the second. (D): Same as (C),
except using empirical food webs from Mangal database, where richness.
The outlier on (D) is a 714 species food-web.}\label{fig:geometric}
}
\end{figure}

\hypertarget{false-negatives-as-a-product-of-relative-abundance}{%
\section{False-negatives as a product of relative
abundance}\label{false-negatives-as-a-product-of-relative-abundance}}

We now show that the realized false-negative rate of species
interactions changes drastically with sampling effort due to the
intrinsic variation of the abundance of individuals of each species
within a community. We do this by simulating the process of observation
of species interactions, applied both to 243 empirical food webs from
the Mangal database (Banville \emph{et al.} 2021) as well as random
food-webs generated using the niche model (Williams \& Martinez 2000), a
simple generative model of food-web structure. Our neutral model of
observation assumes each observed species is drawn from the distribution
in proportion to each species' abundance at that place and time.

The abundance distribution of a community can be reasonably-well
described by a log-normal distribution (Volkov \emph{et al.} 2003). In
addition to the log-normal distribution, we also tested the case where
the abundance distribution is derived from power-law scaling
\(Z^{(log(T_i)-1)}\) where \(T_i\) is the trophic level of species \(i\)
and \(Z\) is a scaling coefficient (Savage \emph{et al.} 2004), which
yields the same qualitative behavior. The practical consequence of
abundance distributions spanning many orders of magnitude of abundance
is that observing two ``rare'' species interacting requires two low
probability events: observing two rare species at the same time.

To simulate the process of observation, for an ecological network \(M\)
with \(S\) species, we sample abundances for each species from a
standard-log-normal distribution. For each true interaction in the
adjacency matrix \(M\) (i.e. \(M_{ij}=1\)) we estimate the probability
of observing both species \(i\) and \(j\) at a given place and time by
simulating \(n\) observations of all individuals of any species, where
the species of the individual observed at the \(\{1,2,\dots,n\}\)-th
observation is drawn from the generated log-normal distribution of
abundances. For each pair of species \((i,j)\), if both \(i\) and \(j\)
are observed within the n-observations, the interaction is tallied as a
true positive if \(M_{ij}=1\). If only one of \(i\) or \(j\) are
observed---but not both---in these \(n\) observations, but \(M_{ij}=1\),
this is counted as a false-negative, and a true-negative otherwise.

In fig.~\ref{fig:geometric}(C) we see this model of observation applied
to niche model networks across varying levels of species richness, and
in fig.~\ref{fig:geometric}(D) the observation model applied to Mangal
food webs. For all niche model simulations in this manuscript, for a
given number of species \(S\) the number of interactions is drawn from
the flexible-links model fit to Mangal data (MacDonald \emph{et al.}
2020), effectively drawing the number of interactions \(L\) for a random
niche model food-web as

\[L \sim  \text{BetaBinomial}(S^2-S+1,\mu\phi, 1-\mu\phi)\]

where the MAP estimate of \((\mu, \phi)\) applied to Mangal data from
{[}{]} MacDonald et al.~2020 is \((\mu=0.086, \phi=24.3)\). All
simulations were done with 500 independent replicates per unique number
of observations n.~All analyses presented here are done in Julia v1.8
(Bezanson \emph{et al.} 2015) using both EcologicalNetworks.jl v0.5 and
Mangal.jl v0.4 (Banville \emph{et al.} 2021) and are hosted at
(GITHUB\_LINK\_TODO). Note that the empirical data, for the reasons
described above, very likely already contains many false negatives,
we'll revisit this issue in the final section.

From fig.~\ref{fig:geometric}(C) it is evident that the number of
species considered in a study is inseparable from the false-negative
rate in that study, and this effect should be taken into account when
designing samples of ecological networks in the future. We see a similar
qualitative pattern in fig.~\ref{fig:geometric}(D) where the FNR drops
off quickly as a function of observation effort, mediated by total
richness. The practical consequence of the bottom row of Figure 1 is
whether the total number of observations of all species (the x-axis) for
the range of possible FNR we deem acceptable (the y-axis) is feasible.
This raises two points: first, empirical data on interactions are
subject to the practical limitations of funding and human-work hours,
and therefore existing data tend to fall on the order of hundreds or
thousands observations of individuals per site. Clear aggregation of
data on sampling effort has proven difficult to find and a meta-analysis
of network data and sampling effort seems both pertinent and necessary,
in addition to the effects of aggregation of interactions across
taxonomic scales (Gauzens \emph{et al.} 2013; Giacomuzzo \& Jordán
2021). This inherent limitation on in-situ sampling means we should
optimize where we sample across space so that for a given number of
samples, we obtain the maximum information possible. Second, what is
meant by ``acceptable'' FNR? This raises the question: does a shifting
FNR lead to rapid transitions in the our ability inference and
predictions about the structure and dynamics of networks, or does it
produce a roughly linear decay in model efficacy? We explore this in the
next section.

We conclude this section by advocating for the use of neutral models
similar to above to generate expectations about the number of
false-negatives in a data set of a given size. This could prove fruitful
both for designing surveys of interactions but also because we may want
to incorporate models of imperfect detection error into predictive
interactions models, as Joseph (2020) does for species occurrence
modeling. Additionally, we emphasize that one must consider the context
for sampling---is the goal to detect a particular species (as in
fig.~\ref{fig:geometric}(C)), or to get a representative sample of
interactions across the species pool? This argument is well-considered
when sampling individual species (Willott 2001), but has not yet been
adopted for designing samples of communities.

\hypertarget{positive-associations-increase-the-false-negative-rate}{%
\section{Positive associations increase the false-negative
rate}\label{positive-associations-increase-the-false-negative-rate}}

This model above doesn't consider the possibility that there are
positive or negative associations which shift the probability of
observing two species together due to their interaction (Cazelles
\emph{et al.} 2016). However, here we demonstrate that the probability
of observing a false negative can be higher if there is some positive
association in the occurrence of species \(A\) and \(B\).If we denote
the probability that we observe the co-occurrence of two species \(A\)
and \(B\) that we know interact as \(P(AB)\) and if there is no
association between the marginal probabilities of observing \(A\) and
observing \(B\), denoted \(P(A)\) and \(P(B)\) respectively, then the
probability of observing their co-occurrence \(P(AB) = P(A)P(B)\). In
the other case where there is some positive strength of association
between observing both \(A\) and \(B\) because this interaction is
``important'' for each species, then the probability of observation both
\(A\) and \(B\), \(P(AB)\), is greater than \(P(A)P(B)\) as \(P(A)\) and
\(P(B)\) are not independent and instead are positively correlated,
i.e.~\(P(AB)> P(A)P(B)\). In this case, the probability of observing a
false negative in our naive model from fig.~\ref{fig:geometric}(A) is
\(p_{fn}= 1-P(AB)\), which due to the above inequality implies
\(p_{fn} >1-P(A)P(B)\). This indicates an increasingly greater
probability of a false negative as \(P(AB) \to P(AB) \gg P(A)P(B)\).
However, this still does not consider variation in species abundancein
space and time (Poisot \emph{et al.} 2015). If positive or negative
associations between species structure variation in the distribution of
\(P(AB)\) across space/time, then the spatial/temporal biases induced by
data collection would further impact the realized false negative rate,
as the probability of false negative would not be constant for each pair
of species across sites.

To test for this association in data we scoured Mangal for datasets with
many spatial or temporal replicates of the same system. For each
dataset, we compute the marginal probability \(P(A)\) of occurrence of
each species \(A\) across all networks in the dataset. For each pair of
interacting species \(A\) and \(B\), we then compute and compare the
probability of co-occurrence if each species occurs independently,
\(P(A)P(B)\), to the empirical joint probability of co-occurrence,
\(P(AB)\). Following our analysis above, if \(P(AB)\) is greater than
\(P(A)P(B)\), then we expect our neutral estimates of the FNR above to
underestimate the realized FNR. In fig.~\ref{fig:mangal}, we see the
difference between \(P(AB)\) and \(P(A)P(B)\) for the seven suitable
datasets with enough spatio-temporal replicates and a shared taxonomic
backbone (meaning all individual networks use common species
identifiers) found on Mangal to perform this analysis. Further details
about each dataset are reported in tbl.~\ref{tbl:id}.

\begin{figure}
\hypertarget{fig:mangal}{%
\centering
\includegraphics{./figures/fig2.png}
\caption{The difference between joint-probability of co-occurrence
(P(AB)) and expected probability of co-occurrence under independence
(P(A)P(B)) for interacting species for each dataset. The red-dashed line
indicates 0. Each histogram represents a density, meaning the area of
the entire curve sums to 1. The continuous density estimate (computed
using local smoothing) is shown in grey. The p-value on each plot is the
result of a one-sided T-test comparing the mean of each distribution to
0.}\label{fig:mangal}
}
\end{figure}
